\racets: Policy{-}Agnostic Programming in Racket

### 2018-07-24 paper/paper.tex
adding related work

>>>>

\section{Related Work and Future Directions}

To the best of our knowledge, we are the first authors to present an
implementation of faceted execution using hygienic macros. There are
several threads of related work in dynamic information flow and
programming paradigms for information flow.

Information-flow was first formalized by \citet{Denning:1976}.
In her seminal work on a lattice model
for information flow, she outlined challenges and potential solutions
to static information-flow checking. Subsequently, \citet{Goguen:82}
defined noninterference, which formalized the idea that privileged
data should not influence publicly observable
outputs. \citet{Clarkson:08} later recognized that
information-flow properties fit into a class of program properties
that could not be characterized by a single trace of a program, but
rather a set of traces, and called these hyperproperties.

Along with definitions of information flow, there has also been
significant interest in mechanisms for enforcing information
flow. This work can be broadly divided into static and dynamic
enforcement mechanisms for information flow security. Of the
mechanisms for static information flow, security type systems have
gained the most use. First introduced by \citet{Volpano:1997}, these
type systems augment the binding environment to track the privilege of
variables and prevent writes to variables that would violate
noninterference. Myers leveraged this idea to produce Jif, a variant
of Java with an information-flow type
system~\cite{Myers:1999}. Security type systems have been subsequently
extended to accommodate concurrent programs~\cite{Zdancewic:03} and
flow sensitivity~\cite{Hunt:2006}. Faceted execution does not require
annotating the program with security types, but at the expense of
losing a static characterization of the program's security in its type
system.

\citet{Devriese:10} first introduced secure multi-execution as a
dynamic enforcement technique for information flow. Secure
multi-execution runs $2^k$ copies of a program in parallel, where each
run represents a subset of $\wp(\var{Prin})$, where $\var{Prin}$ is a
set of principals. For example, if the principals in the program are
Alice and Bob, secure multi-execution executes four copies of the
program: one that replaces all secret inputs by $\bot$, one that
replaces Bob's input by $\bot$ but Alice's input by the true input,
one for Bob's input, and one with access to all privilaged
information. When external effects are made (e.g., writing to disc),
the runtime can select which variant to use based on a policy. Secure
multi-execution prevents information flow violations at runtime by
ensuring that observations which violate the information-flow policy
receive a view of the data computed without access to the secret
inputs. Secure multi-execution has been extended in a variety of ways,
e.g., scaling to its implementation in web browsers~\cite{Bielova:11},
adding declassification in a granular way~\cite{Rafnsson:13}, and even
preventing side-channel attacks~\cite{Kashyap:11}.

As the number of principals increases, secure multi-execution's
overhead increases exponentially, unnecessarily duplicating work not
influenced by secret inputs. Austin et al. introduced faceted
execution as an optimization of secure multi-execution
in~\cite{Austin:2012}. Instead of treating the whole program as a
potentially-secret computation, faceted execution realizes that
influence can be tracked and propagated in a granular way using
facets. Notably, Austin et al.'s work does not include first-class
labels, as it was simulating secure multi-execution, where the
principals could not be dynamically generated.

At the same time, Yang et al. first implemented Jeeves, a language
allowing policy-agnostic programming~\cite{Yang:2012}. Policy-agnostic
programming takes the view that programs should be written without
regard to a particular privacy policy, because as the policy changes,
correctly updating program logic is cumbersome and
error-prone. Policy-agnostic programming was first implemented in the
domain-specific language Jeeves, using an SMT solver to decide which
view of secret data to reveal based on a policy. Later, both authors
collaborated to implement Jeeves using faceted
execution.~\cite{Austin:2013}. This formulation includes first-class
labels, and is the basis for our concrete semantics.

Several other efforts into dynamic analysis for information flow are
worth noting. \citet{Stefan:2011} first presented \code{LIO}---a monad
(with implementation in Haskell) that tracks privilege of the current
program counter and forbids effects that would violate the security
policy. It might be surprising that \code{LIO}
works well for Haskell programs, given that faceted execution is
more precise than \code{LIO}---allowing values to become faceted
rather than halting the program. One key difference is that Haskell
programs emphasize purity while languages such as
JavaScript (the original target of faceted execution) does not, so much
of the machinery for faceted execution's effect on the store is less
interesting. Several authors have implemented related systems to
\code{LIO}, including variants of faceted execution \cite{Schmitz:2016}
and variants of \code{LIO} that extend its
power to arbitrary monad transformers~\cite{james}. We believe that it
would be possible to implement a variant of our technique that would
give similar insights to programs using \code{LIO}, though much of the
interesting machinery for handling state may be unnecessary.
